Partitioned Local Depth (PaLD) Community Analyses in R

### 2023-05-02 r-journal/dagostino-mcgowan.Rmd
update

Cohesion is particularly useful when considering data with varying local density, see discussion in \citet{berenhaut2022social}.  Note that the PaLD algorithm is able to detect the eight natural groups within the data without the use of any additional inputs (e.g., number of clusters) nor optimization criteria.  Despite the user input of the "correct" number of clusters (i.e., $k = 8$) both *k*-means and hierarchical clustering did not give the desired result. For further examples, discussion, and theoretical results, see \citet{berenhaut2022social}.

>>>>

Cohesion is particularly useful when considering data with varying local density; see \citet{berenhaut2022social} for further examples, discussion, and theoretical results.  Note that the PaLD algorithm is able to detect the eight natural groups within the data without the use of any additional inputs (e.g., number of clusters) nor optimization criteria.  Despite the user input of the "correct" number of clusters (i.e., $k = 8$) both *k*-means and hierarchical clustering did not give the desired result.

### Cultural and Pyschological distance analysis

**TODO intro to data here** \citep{muthukrishna2020beyond}

First, we will create the cohesion matrix using the `cultures` data, provided in \CRANpkg{pald}, and proceed to plot the community structure.

```{r}
cultures_cohesion <- cohesion_matrix(cultures)
```

```{r, eval = FALSE}
plot_community_graphs(
  cultures_cohesion,
  edge_width_factor = 30,
  emph_strong = 3,
  vertex.label.cex = 0.7,
  vertex.size = 3,
  vertex.label.dist = 1
)
```

```{r, include = FALSE, results = "hide"}
set.seed(7)
png("fig7.png", res = 300, height = 3000, width = 3000)
plot_community_graphs(
  cultures_cohesion,
  edge_width_factor = 30,
  emph_strong = 3,
  vertex.label.cex = 0.7,
  vertex.size = 3,
  vertex.label.dist = 1
)
dev.off()
```

```{r figculture, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Community structure for cultural distance data.", out.width = "100%"}
knitr::include_graphics("fig7.png")
```

This example demonstrates key features of the PaLD approach that distinguish it from commonly used clustering techniques. In addition to viewing the community structure as seen in Figure \ref{fig:figculture}, we can also visualize a two dimensional plot of cohesion versus distance for each of the included distance pairs using the `dist_cohesion_plot()` function, as seen below (Figure \ref{fig:figco}).

```{r figco, fig.cap = "Distance versus cohesion for each TODO The identified communties are shown via the 5 colors -- these map to the same clusters idenfitied in Figure 7."}
dist_cohesion_plot(cultures)
```

Notice here the magnitude of the distances within each of the identified communities varies substantially between regions, yet the cohesion remains strong. In other words, focusing particularly on identified community consisting of regions in the United States (light purple) compared to the community consisting of regions in India (turquoise) the strong cohesions are similar (Figure \ref{fig:figculture}), yet the distances themselves are markedly different (Figure \ref{fig:figco}). In fact, the largest distance between regions in the Unites States is smaller than that between any two regions in India (Figure \ref{fig:figco}). Classic distance-based clustering techniques tend to be unable to maintain this distinction often breaking India into several smaller communities. Additionally, currently available techniques require specification of parameters, as seen in the previous section. For further discussion of the cultural distance data in relation to community analysis see \citet{berenhaut2022social}.

### 2023-05-02 r-journal/dagostino-mcgowan.tex
update

\begin{Schunk}
\begin{Sinput}
graphs <- community_graphs(cohesion)
graphs[["G_strong"]]
\end{Sinput}
\begin{Soutput}
#> IGRAPH e4c9a8b UNW- 6 3 -- 
#> + attr: name (v/c), weight (e/n)
#> + edges from e4c9a8b (vertex names):
#> [1] A--B A--E D--F
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
french_neighbors <- igraph::neighbors(cognate_graph_strong, "French")
french_neighbors
\end{Sinput}
\begin{Soutput}
#> + 8/87 vertices, named, from 81f58d9:
#> [1] Italian         Ladin           Provencal       Walloon        
#> [5] French_Creole_C French_Creole_D Spanish         Catalan
\end{Soutput}
\end{Schunk}

@@

Cohesion is particularly useful when considering data with varying local
density, see discussion in \citet{berenhaut2022social}. Note that the
PaLD algorithm is able to detect the eight natural groups within the
data without the use of any additional inputs (e.g., number of clusters)
nor optimization criteria. Despite the user input of the ``correct''
number of clusters (i.e., \(k = 8\)) both \emph{k}-means and
hierarchical clustering did not give the desired result. For further
examples, discussion, and theoretical results, see
\citet{berenhaut2022social}.

>>>>

\begin{Schunk}
\begin{Sinput}
graphs <- community_graphs(cohesion)
graphs[["G_strong"]]
\end{Sinput}
\begin{Soutput}
#> IGRAPH 2f2c953 UNW- 6 3 -- 
#> + attr: name (v/c), weight (e/n)
#> + edges from 2f2c953 (vertex names):
#> [1] A--B A--E D--F
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
french_neighbors <- igraph::neighbors(cognate_graph_strong, "French")
french_neighbors
\end{Sinput}
\begin{Soutput}
#> + 8/87 vertices, named, from 828fab3:
#> [1] Italian         Ladin           Provencal       Walloon        
#> [5] French_Creole_C French_Creole_D Spanish         Catalan
\end{Soutput}
\end{Schunk}

@@

Cohesion is particularly useful when considering data with varying local
density; see \citet{berenhaut2022social} for further examples,
discussion, and theoretical results. Note that the PaLD algorithm is
able to detect the eight natural groups within the data without the use
of any additional inputs (e.g., number of clusters) nor optimization
criteria. Despite the user input of the ``correct'' number of clusters
(i.e., \(k = 8\)) both \emph{k}-means and hierarchical clustering did
not give the desired result.

\hypertarget{cultural-and-pyschological-distance-analysis}{%
\subsubsection{Cultural and Pyschological distance
analysis}\label{cultural-and-pyschological-distance-analysis}}

\textbf{TODO intro to data here} \citep{muthukrishna2020beyond}

First, we will create the cohesion matrix using the \texttt{cultures}
data, provided in \CRANpkg{pald}, and proceed to plot the community
structure.

\begin{Schunk}
\begin{Sinput}
cultures_cohesion <- cohesion_matrix(cultures)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
plot_community_graphs(
  cultures_cohesion,
  edge_width_factor = 30,
  emph_strong = 3,
  vertex.label.cex = 0.7,
  vertex.size = 3,
  vertex.label.dist = 1
)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{figure}
\includegraphics[width=1\linewidth]{fig7} \caption[Community structure for cultural distance data]{Community structure for cultural distance data.}\label{fig:figculture}
\end{figure}
\end{Schunk}

This example demonstrates key features of the PaLD approach that
distinguish it from commonly used clustering techniques. In addition to
viewing the community structure as seen in Figure \ref{fig:figculture},
we can also visualize a two dimensional plot of cohesion versus distance
for each of the included distance pairs using the
\texttt{dist\_cohesion\_plot()} function, as seen below (Figure
\ref{fig:figco}).

\begin{Schunk}
\begin{Sinput}
dist_cohesion_plot(cultures)
\end{Sinput}
\begin{figure}
\includegraphics{dagostino-mcgowan_files/figure-latex/figco-1} \caption[Distance versus cohesion for each TODO The identified communties are shown via the 5 colors -- these map to the same clusters idenfitied in Figure 7]{Distance versus cohesion for each TODO The identified communties are shown via the 5 colors -- these map to the same clusters idenfitied in Figure 7.}\label{fig:figco}
\end{figure}
\end{Schunk}

Notice here the magnitude of the distances within each of the identified
communities varies substantially between regions, yet the cohesion
remains strong. In other words, focusing particularly on identified
community consisting of regions in the United States (light purple)
compared to the community consisting of regions in India (turquoise) the
strong cohesions are similar (Figure \ref{fig:figculture}), yet the
distances themselves are markedly different (Figure \ref{fig:figco}). In
fact, the largest distance between regions in the Unites States is
smaller than that between any two regions in India (Figure
\ref{fig:figco}). Classic distance-based clustering techniques tend to
be unable to maintain this distinction often breaking India into several
smaller communities. Additionally, currently available techniques
require specification of parameters, as seen in the previous section.
For further discussion of the cultural distance data in relation to
community analysis see \citet{berenhaut2022social}.
